---
title: "Group Project Plan"
author: "M. Lewis, P. Magee, M. Willis, and E. Wong"
format: pdf
editor: visual
bibliography: citations.bib
csl: chicago-author-date.csl
---

```{r, include = FALSE}
library(tidyverse)
```

# Project Title:

Predicting the Forgiveness of Paycheck Protection Program (PPP) Loans in Washington, DC.

# Group Members

-   Michael Lewis (427)

-   Peter Magee (627)

-   Margaret Willis (627)

-   Ethan Wong (427)

# Topic

Using publicly available data on PPP loans for business in DC, we seek to develop models that will explore/predict what factors were important for loan forgiveness -- and if so by how much -- and whether these things can be used to predict whether a loan was forgiven.

Below is what our dataset looks like.

```{r echo=FALSE, message=FALSE, warning=FALSE}
read_csv("PPP details.csv") -> dc
dc %>% head()
```

# Questions of Interest

Throughout the COVID-19 pandemic, an estimated \$525 Billion USD was disbursed as part of the federal government's Paycheck Protection Program (PPP) @house_fact_2022. These loans, while crucial for protecting American industry and the workforce, have been met with allegations of fraud and controversy -- in some cases spurring litigation @pfeiffer_how_2023. Today, the vast majority of the loans have been forgiven @pfeiffer_how_2023 @noauthor_firms_2021. Analyzing this load data presents an opportunity to better understand (1) what business characteristics are associated with higher loan amounts, (2) what factors led to loan forgiveness -- and whether the amount forgiven varied based on these things, and finally (3) whether ML techniques can be used to predict whether a loan was forgiven off these considerations.

### Regression:

-   How can we assess the relationship between `Amount forgiven` and other variables like `Rent`, `Jobs reported`, `Business type`, `Industry`, `Age of business`, `Utilities`, etc.?

### Classification:

-   What methods have the lowest MSE for the correct prediction of whether loans were forgiven or not.

# Approaches

-   We propose using logistic regression, KNN, LDA, QDA to predict whether a loan was forgiven in whole or not.
-   We also propose using linear regression (with cross-validation) to quantify the numerical effect of business characteristics on amount forgiven.

# Group Responsibilities

-   Tidy and clean data prior to analysis (Michael and Pete).

-   Serve as point of contact for reaching out to Professor Ressler (Margaret)

-   Formatting poster (assuming we wish to go that route).

-   We will each select an SML strategy.

# Deliverable

-   Our final deliverable will be in the poster format.

# References
